\section{Importance Functions for Variance Reduction}
\label{sec:Importance}

The effective use of variance reduction techniques can lead to a faster time to
a desired solution and a reduced variance in the specified tally. However,
specifying variance reduction parameters is not always a straightforward
procedure. In simple geometries, a user might intuitively understand which
regions of a problem may contribute more to a desired solution, but for more
complex geometries, this may not be so obvious.
In the following subsections, the theory in determining which regions of a
problem are important to eliciting a tally response will be described.
The first topic discussed will be
the concept of importance and obtaining a measure of importance with
Monte Carlo sampling.
Second, the adjoint equation and its relation to importance will be introduced.
Last, the contributon solution and how its relation to tally responses is
reviewed.

\subsection{The Concept of Importance}
\label{sec:Importance}

The concept of importance is, in essence, a means of defining which regions of
a problem that are likely to contribute to a
response and which are less likely to contribute to a response. The regions
that are more likely to generate a response will have a
higher importance than those that do not. As one might imagine, this may be
intuitively deduced for a simple problem, but in a more complex problem
that might not be possible.
If an importance
function for a system can be obtained computationally, then that importance
function can be
strategically used in variance reduction techniques to speed up the Monte Carlo
calculations.

As described in Section \ref{subsec:AutomatedMCVR}, Booth
\cite{booth_automatic_1982} proposed
a method to quantify a cell's importance within a Monte Carlo simulation
(Eq. \eqref{eq:BoothImp}). In
this method, Booth suggested estimating the cell's importance using Monte Carlo
transport as:
\begin{equation*}
  Importance  = \frac{\text{score of particles leaving the cell}}
                     {\text{weight leaving the cell}}.
\end{equation*}
This particular calculation of importance
follows from the intuitive explanation for importance in the preceding
paragraph. Recall from Section \ref{subsec:MCVR} that in variance reduction
methods, the population of particles is increased in important regions such that
the number of samples or particles contributing to a tally increases, but the
total problem weight is conserved. More important regions should have many
lower-weight particles to reduce the tally variance. Using Booth's bookkeeping
method for estimating regional importance,
if a cell has a greater weight leaving the cell than the number
of particles, that means that the relative contribution of that cell to the
tally is likely to be lower than other regions. If, instead, the number of
particles leaving the cell is greater than the weight leaving the cell, then
that region is more important to the tally response, because that particle
population is higher than other cells.

While this
estimation of the importance requires only a Monte Carlo forward calculation of
the problem, it was referred to as the forward-adjoint importance generator
\cite{booth_automatic_1982, booth_deep_1982, booth_importance_1984}, because the
bookeeping tracked by Eq. \eqref{eq:BoothImp} was a forward-approximation of the
adjoint. Adjoint theory and how it relates to importance will be discussed in
Section
\ref{sec:AdjointImportance}.
Booth's estimation of importance was used to generate
weight window target values inversely proportional to the importance.
In this case, the track weight times the expected score is approximately
constant in the problem. Choosing this inverse relationship between the weight
window and importance is common practice in variance reduction, and is often a
good choice because it is nearly optimal over a broad range of the problem
phase-space \cite{booth_common_2012}.

It should be noted that Booth's method was reliant on the statistical
precision of the cells sampled to generate their importances. For
deep-penetration problems, obtaining a ``good'' estimate of the cell importances
many mean free paths from the forward source took several iterations. With
large fluctuations between iterations, this had the potential to
be a very slow and
computationally inefficient way to calculate importance in a problem. Using a
solution of the adjoint that is equally valid across all of the problem space
would be more ideal for deep-penetration problems.

\subsection{The Adjoint Solution for Importance}
\label{sec:AdjointImportance}

Using the solution of the adjoint formulation of the neutron transport
equation is one of the most widely recognized methods for generating
an importance function. This subsection will begin with a brief summary of
adjoint theory. A discussion on how
the adjoint solution differs physically from the forward solution for a
particular problem will follow. Last, some early investigations on how the
adjoint
and importance are related will be summarized.

\subsubsection{Theory}

In previous sections we have reviewed the statistical precision of Monte
Carlo-based methods, and how sampled results from Monte Carlo can be sped up
with variance reduction methods. We have also briefly addressed the forward- and
the adjoint- solutions for a particular problem. In neutron transport, the
integral form of the forward, steady-state, particle transport equation can be
defined as:
\begin{multline}
\hat\Omega \cdot \nabla \psi
        (\vec {r} ,E,\:\hat\Omega)+\Sigma _{ t }
        (\vec{r},E)\psi (\vec { r } ,E,\:\hat\Omega) = \\
        \int _{ 4\pi  } \int _{ 0 }^{ \infty  } \Sigma _{ s }(E'\rightarrow E,
        \hat\Omega'\rightarrow\hat\Omega)\psi (\vec { r } ,E',\: \hat\Omega')dE'
        \:d\hat\Omega' + q_{e}(\vec { r } ,E, \:\hat\Omega),
\label{eq:F-NTE}
\end{multline}
where $\vec { r }$, $E$, and $\hat\Omega$, are the direction, energy, and angle
independent variables totalling seven dimensions in total,
$\psi$ is the neutron flux, $\Sigma$ is the neutron
interaction (scattering, absorption, total) cross section, and $q_{e}$ is the
external fixed source. Alternatively, this can be written in operator form,
\begin{equation}
  H \psi = q_{e}
\label{eq:F-NTE2}
\end{equation}
where $H$ represents the streaming, scattering, and absorptive terms from Eq.
\eqref{eq:F-NTE}, $\psi$ is the angular flux as it is in Eq. \eqref{eq:F-NTE}, and
$q_{e}$ is the source term.

The forward transport equation tells us where particles are moving
throughout the system that we define by equation \ref{eq:F-NTE}. Of note, the
particles move in the scattering term from $E'$ into $E$, and from $\hat\Omega'$
into $\hat\Omega$. Therefore for a particular problem with a given $q_{e}$,
particles start at $q_e$ and move throughout the system,
either downscattering in energy, streaming out of the problem,
are absorbed by the problem materials, or
induce a response at the tally location.

The adjoint equation of the same form, in contrast, can be expressed as:
\begin{multline}
-\hat\Omega \cdot \nabla \psi^{\dagger}
        (\vec {r} ,E,\:\hat\Omega)+\Sigma _{ t }
        (\vec{r},E)\psi^{\dagger}  (\vec { r } ,E,\:\hat\Omega)
       = \\
        \int _{ 4\pi  } \int _{ 0 }^{ \infty  } \Sigma _{ s }(E\rightarrow E',
        \hat\Omega\rightarrow\hat\Omega')\psi^{\dagger}  (\vec { r } ,E',\:
        \hat\Omega')dE' \:d\hat\Omega' + q_{e}^\dagger(\vec { r } ,E,
        \:\hat\Omega),
\label{eq:A-NTE}
\end{multline}
or in operator form as
\begin{equation}
  H^{\dagger} \psi^{\dagger} = q_{e}^{\dagger}.
\label{eq:A-NTE2}
\end{equation}
Note here that the particles in the adjoint equation move from $E$ into $E'$, and
from $\hat\Omega$ into $\hat\Omega'$, which indicates an upscattering in energy
and a reversal of direction when compared to the forward problem.
The external source, too, is different, changing
from $q_{e}$ to $q_{e}^\dagger$.

To solve the adjoint problem the adjoint source, $q_{e}^{\dagger}$,
can be chosen such that it has the potential to reveal information about the
forward problem. In Monte Carlo variance reduction, we seek to obtain
information on the
detector or tally response for the system.
The response for the forward problem
given a defined source distribution  $q(\vec{r}, E, \hat{\Omega})$ is
\begin{subequations}
\begin{equation}
  R_{tally} = \int_{4\pi} \int_{V} \int_{E} \psi(\vec{r}, E, \hat{\Omega})
  \Sigma_{tally}(\vec{r},E, \hat\Omega) dE dV d\Omega ,
\end{equation}
which can be simplified using bracket notation, where the brackets indicate an
integration over all phase-space,
\begin{equation}
  R_{tally} = \langle \psi \Sigma_{tally} \rangle .
\end{equation}
\end{subequations}
$\psi$ is the angular flux and $\Sigma_{tally}$ is the effective tally
response function. For a simple source-detector problem, we choose
$q_{e}^{\dagger}$ to be $\Sigma_{tally}$, or that the adjoint source is the
tally/detector response function
for the system. Therefore the adjoint particles start at low energy at the detector
location, move away from the adjoint source (the detector location), and scatter
up in energy. By making the choice that $q_{e}^{\dagger} = \Sigma_{tally}$, the
response function can be written as a product for the forward flux and the
adjoint source
\begin{equation}
  R_{tally} = \langle \psi q^{\dagger} \rangle .
  \label{eq:response1}
\end{equation}
By using the adjoint identity and the same operators $H$ from Eqs. \eqref{eq:F-NTE2}
and \eqref{eq:A-NTE2}
\begin{equation}
  \langle \psi, H^{\dagger} \psi^{\dagger} \rangle =
  \langle \psi^{\dagger}, H \psi \rangle ,
\end{equation}
Equation \eqref{eq:response1} can also be written to be a function of the
adjoint flux and the forward source distribution
\begin{equation}
  R = \langle \psi^{\dagger} q \rangle .
  \label{eq:response2}
\end{equation}

At this point, we know that the solution to the adjoint problem has
transported particles from the adjoint source (which is the detector or tally
location) into the problem phase-space. The adjoint particles are
upscattered in energy and are transported in $-\Omega$ relative to the forward
problem. However, it may not be immediately obvious how this adjoint solution
relates to importance for the forward solution. Let us start with a simple
illustrative example: a monoenergetic, monodirectional, point source. The
forward source takes the form of a delta function:
\begin{equation*}
  q(\vec{r}, E, \hat{\Omega}) = \delta(\vec{r}-\vec{r}_0) \delta(E-E_0)
  \delta(\hat{\Omega}-\hat{\Omega}_0) .
\end{equation*}
Using this definition of the forward source and evaluating Eq.
\eqref{eq:response2}, we obtain
\begin{equation*}
  \begin{split}
    R &= \langle \psi^{\dagger} q \rangle \\
    &= \int_{V} \int_{E} \int_{\Omega} \psi^{\dagger}(\vec{r}, E, \hat{\Omega})
       q(\vec{r}, E, \hat{\Omega}) d\Omega dE dV \\
       & = \psi^{\dagger}(\vec{r}_0, E_0, \hat{\Omega}_0).
\end{split}
\end{equation*}
This result shows that the solution to the adjoint equation is the detector
response for the forward problem. As a result, the adjoint flux can be used as
an indicator of a particle produced in $\vec{r}, E, \hat{\Omega}$ contributing
to a response in the system. This indicator can be thought of as the particle's
importance to achieving the tally or response objective. Consequently, it is
often said that the adjoint is the importance function for the problem.

The adjoint solution is used in nuclear engineering for a number of
applications, including reactor physics and perturbation theory
\cite{lewins_importance_1965, lewins_developments_1968,
greenspan_developments_1976, lux_monte_1991}.
However, Goertzel and Kalos' early work recognized its
application for deep-penetration radiation shielding applications.
Goertzel and Kalos \cite{goertzel_monte_1958}
showed analytically
that the exact adjoint solution, if used as an importance or weighting
function for the forward Monte Carlo calculation, will result in a zero variance
solution for the forward Monte Carlo problem. Further, Kalos
\cite{kalos_importance_1963} showed in a 1D infinite hydrogen slab problem that
an analytically-derived adjoint importance function significantly helped neutron
transport for deep-penetration problems.

Goertzel and Kalos' finding that an exact adjoint can lead to a zero variance
solution means that if a single particle is transported with the adjoint
weighting function, its score will be the same as the total system response.
Only a single particle is required to get an exact solution for the forward
problem. This is prohibitive because obtaining
an exact adjoint solution is just as computationally expensive as getting
an exact forward solution. As a result, one seeks to obtain a good estimate of the
adjoint solution based on computational limitations. A good importance estimate
should help reduce the variance in a reasonable amount of time, and be
relatively computationally inexpensive. A Monte Carlo solution can
provide a continuous solution over the problem phase-space. However, as
discussed in Section \ref{subsec:MCVR}, the quality of this adjoint solution
relies on the number of samples used to calculate it and that may take a
significant amount of time. A deterministic solution
offers equal solution confidence across the entire problem, but it is
discretized
in space, energy, and angle. For deep-penetration importance functions, we will
opt for deterministically-obtained solutions due to the solution's equally
distributed validity.

\subsubsection{Implementation}

Coveyou, Cain and Yost \cite{coveyou_adjoint_1967} expanded on Goertzel and
Kalos' work by interpreting in which ways the adjoint solution could be adapted
for Monte Carlo variance reduction. In particular, they investigated the choice
of biasing schemes and how effective they were at variance reduction for a
simple one-dimensional problem. They reiterated that the adjoint solution is a
good estimate for importance, but should not be calculated explicitly, and
rather estimated by a simpler model. The adjoint function is not necessarily the
most optimal importance function; however, it is likely the closest and most
obtainable estimate of importance that can be calculated
\cite{coveyou_adjoint_1967}. They
concluded that source biasing by the solution to the adjoint equation or by the
expected response is an best choice for Monte Carlo variance reduction, as it
can be used independently of any other variance reduction technique, and
provides good results.

Tang and Hoffman \cite{tang_monte_1988} built off of the parameters derived by
Coveyou et al \cite{coveyou_adjoint_1967} to generate variance reduction
parameters automatically for fuel cask dose rate analyses. In their work, Tang
and Hoffman utilized a one-dimensional discrete ordinates code XSDRNPM-S to
calculate the adjoint fluxes for their shielding problems. The results from this
calculation were then used to generate biasing parameters for Monte Carlo;
specifically they aimed at generating parameters for energy biasing, source
biasing, Russian roulette and splitting, and next event estimation
probabilities. They implemented their work in the SAS4 module in SCALE
\cite{SCALE6_1}; it was
one of the earlier implementations of a fully-automated deterministic
biasing procedure for Monte Carlo.

\subsection{The Contributon Solution for Importance}
\label{sec:ContributonImportance}

Contributon theory is another useful metric that can be used as a measure of
importance
\cite{williams_generalized_1991,williams_contributorn_1992,williams_contributon_study}.
However, contributon theory quantifies importance differently than
adjoint theory. In contributon transport, a pseudo-particle, the
texit{contributon}, is defined. The contributon carries response in the problem
system from the radiation source to a dectector. The
total number of contributons in a system are conserved by the textit{contributon
conservation principle}: all contributons that are emitted from the source
eventually arrive at the detector. Much of the work in this realm has been done
by Williams and collaborators
\cite{williams_generalized_1991,williams_contributorn_1992,williams_contributon_study}.
Williams noted that one could
go so far as to track contributons, rather than real particles in Monte Carlo.
Because every particle transported would eventually reach the detector, this
would lead to a zero variance solution. However, the nature of solving the
contributon equation with Monte Carlo (or any other transport mechanism)
involves knowing the exact solution to the adjoint equation, and so relies on
the same computational obstacles as solving the adjoint transport equation and
using that alone for an importance measure.

The contributon transport equation can be derived in a form analogous
to the forward (Eq. \ref{eq:F-NTE} ) and adjoint (Eq. \ref{eq:A-NTE}) equations.
The derivation of Eq. \ref{eq:Cont-NTE} and its corresponding variables is
available in a number of the sources referenced in this section, so we will
abstain from doing so at this time.
The angular contributon flux is defined
as the product of the forward- and
adjoint- angular- fluxes:
\begin{equation}
\Psi (\vec {r} ,E,\:\hat\Omega) = \psi^{\dagger} (\vec {r} ,E,\:\hat\Omega)
        \psi(\vec {r} ,E,\:\hat\Omega)
\label{eq.Cont-Flux}
\end{equation}
The contributon transport equation is:
\begin{multline}
\hat\Omega \cdot \nabla \Psi (\vec {r} ,E,\:\hat\Omega)
+\widetilde{\Sigma} _{ t }(\vec{r},E,\:\hat\Omega)\Psi (\vec { r } ,E,\:\hat\Omega)
     = \\
        \int _{ 4\pi  } \int _{ 0 }^{ \infty  }
        \widetilde{p}(\vec{r}, \hat\Omega', E'\rightarrow\hat\Omega, E)
        \widetilde{P}(\vec{r}, \hat\Omega',E')
        \widetilde{\Sigma} _{ t }(\vec{r}, E', \hat\Omega')
        \Psi (\vec { r } ,E',\: \hat\Omega')dE' \:d\hat\Omega'
        + \hat p(\vec { r } ,E, \:\hat\Omega) R ,
\label{eq:Cont-NTE}
\end{multline}
where the effective cross sections are given by:
\begin{equation}
\begin{aligned}
\widetilde{\Sigma}_{t}(\vec{r}, E, \hat\Omega) &=
        \widetilde{\Sigma}_{s}(\vec{r}, E, \hat\Omega) +
        \widetilde{\Sigma}_{a}(\vec{r}, E, \hat\Omega)    \\
     &= \frac{\iint \Sigma_{s}(\vec{r},\hat\Omega'\cdot\hat\Omega'',
        E'\rightarrow E'')}{\psi^{\dagger}(\vec{r}, E, \hat\Omega)}
        + \frac{Q^{\dagger}(\vec{r}, E, \hat\Omega)}
        {\psi^{\dagger}(\vec{r}, E, \hat\Omega)}.
\end{aligned}
\end{equation}
The scattering probability of a contributon at position $\vec{r}$, $E'$, and
$\hat\Omega'$ is:
\begin{equation}
\widetilde{P}(\vec{r}, \hat\Omega',E') =
         \frac{\widetilde{\Sigma} _{ s }(\vec{r}, E', \hat\Omega')}
       {\widetilde{\Sigma} _{ t }(\vec{r}, E'Q, \hat\Omega')}
\end{equation}
and
\begin{equation}
\widetilde{p}(\vec{r}, \hat\Omega', E'\rightarrow\hat\Omega, E) =
       \frac{\Sigma_{s}(\vec{r},\hat\Omega'\cdot\hat\Omega,E'\rightarrow E)
       \psi^{\dagger} (\vec{r}, E, \hat\Omega)}
       {\iint \Sigma_{s}(\vec{r},\hat\Omega'\cdot\hat\Omega'',E'\rightarrow
       E'')\psi^{\dagger} (\vec{r}, E'', \hat\Omega'')d\hat\Omega'' dE''}
\end{equation}
is the probability that a contributon scattering at $\vec{r}$, $E'$,
and $\hat\Omega'$ will scatter into $d\hat\Omega$ $dE$.
\begin{equation}
\hat p(\vec{r}, E, \hat\Omega) =
\frac{psi^{\dagger}(\vec{r}, E, \hat\Omega) Q(\vec{r},E,\hat\Omega)}
     {\int \int \int \psi^{\dagger}(\vec{r'},E',\hat\Omega')
     Q(\vec{r'},E',\hat\Omega') d\hat\Omega' dE' dV'}
\end{equation}
is the contributon source distribution.
\begin{equation}
R = \int \int \int \psi^{\dagger}(\vec{r},E,\hat\Omega)Q(\vec{r},E,\hat\Omega)
    d\hat\Omega dE dV
\end{equation}
is the reaction rate.

As mentioned in the previous section, the adjoint flux is an indicator of a
particle's importance to inducing a response. Conversely, the contributon
flux describes what the importance of a particle is to the solution.
Becker's thesis \cite{becker_hybrid_2009} aptly points out that this is
illustrated most dramatically in a source-detector problem, where the forward
source has little importance to the adjoint source, but does have importance to
the problem solution. As a result, both the contributon solution and the adjoint
solution can be considered importance functions for a problem, but the
importance that they describe differs.

Williams recognized the applications of the pseudo-particles, contributons, to
shield design and optimization. In particular, Williams noted that variables
relevant to contributon response were useful in determining transport paths
through media \cite{williams_contributon_study, williams_SCC_shielding}. A study
of the contributon density in a system could enlighten the user on locations
where particles would preferentially be transported, and so designers could
intelligently choose where to place detectors and material. In this way, the
user could find the particles most influential to the response of the system. A
region with high response is the most important to a detector tally. The
variables of response described by Williams are the response potential, the
response current, and the response vorticity \cite{williams_contributorn_1992}.

Contributon theory and spatial channel theory have been applied successfully to
shielding analyses \cite{seydaliev_contributon_2008, williams_SCC_shielding} due
to its ability to show particle flow between a source and response effectively.
Williams and Engle showed that
spatial channel theory can be used in reactor shielding analyses.
In their work, they used
contributon currents to determine preferential flow paths through the Fast Flux
Test Facility (FFTF) \cite{williams_SCC_shielding}.
Seydaliev \cite{seydaliev_contributon_2008} used angle-dependent forward and
adjoint
fluxes and currents to visualize the contributon flux for
simple source-detector problems. In this work, he showed that contributon flow
in the system behaves much like a fluid between the source and detector,
following preferential flow paths more densely.
Seydaliev also observed ray effects in the
contributon flux for high energy photons, and traditional methods like using a
first collision source, did not remedy the issue. The contributon formulation of
particle transport can show important particle flow paths between a source and a
detector, but it is still not immune to computational obstacles that exist for
standard forward- and adjoint- transport.

The past few subsections have described the different means by which importance
can be defined or quantified for a problem.
As discussed in Section \ref{sec:Importance}, generating an importance function
with Monte Carlo is limited in that the quality of the importance map is only as
good as the regions that are sampled. For deep-penetration problems, it may be
prohibitively difficult to obtain adequate importance sampling with traditional
Monte Carlo methods. Deterministically-obtained importance functions
offer the benefit of a solution that is equally valid across all of the problem
solution-space. This is because the deterministic solution's precision is
limited to convergence criteria, not sampling of individual particles. Using a
deterministic solution is often faster and much less computationally-intensive
than Monte Carlo for importance quantification.
As a result, many hybrid methods opt
to use deterministically-obtained importance functions to generate variance
reduction parameters for Monte Carlo transport.
