% (This file is included by thesis.tex; you do not latex it by itself.)

\begin{acknowledgements}

I have been profoundly lucky to have met and worked with some truly inspiring
people throughout my studies in nuclear engineering. Before I begin, I must say
that I would not be here without the contributions--whether large or small--from
all of these wonderful people. I say this with the deepest sincereity: thank you
to all of you.

First, I must thank my advisor, Prof. Rachel Slaybaugh. Rachel, you saw me through
the highest peaks and deepest valleys of my journey in
graduate school, and you provided
nothing but support and encouragement. Thank you for sincerely caring about my
well-being, facilitating my pursuit of knowledge, and enabling me to
push my myself even when I thought it wasn't possible.
Your mentorship has led to my foray into the world
of hybrid methods,
a field to which I never thought I could contribute. Your technical
knowledge and sincere commitment to your students is unparalleled.

I am immensely grateful for my collaborators at Oak Ridge National Laboratory
for their guidance and support. Without their help and mentorship,
my methods development and
software contributions would be far more paltry. Drs. Tara Pandya, Seth Johnson,
Steven Hamilton, and Tom Evans have all been crucial in my professional
transformation, each in their own way.
Tara, the coolest sparkle-transport-developer in the world, thank you for taking
the time to mentor me and help me to learn a topic which was intitially very
overwhelming. Thank you also for your thorough weekly feedback on this project,
which only made this work stronger.
Seth, I aspire to have even a small fraction of
the astronomical quantity of knowledge of Python, software devleopment, and
software usability in your head. Thank you for helping me learn so much more
about methods development than I thoguht possible.
Steven, your cynicism beyond your years is truly astounding. Thank you for the
math help and absurd humor to brighten my days.

I'd also like to thank my remaining committee members,
Profs. Massimiliano Fratoni and John Harte
for serving on my committee and providing feedback on the work contained in this
dissertation.

I have been fortunate to have many great teachers throughout my young career in
research thus far. Thank you Prof. Rick Norman for being a wonderful mentor and
embodiment of a quality teacher. I also have deep gratitude for my undergraduate
advisor, Dr. Todd Palmer, and mentor, Dr. Steve Reese, who gave me the
opportunity to pursue research as a young and inexperienced undergraduate
student. Their
passion for research, wealth of knowledge, and kidness
influenced my choice to go to graduate school.

My pursuits in graduate school would have been far more difficult had my friends
not kindly supported, encouraged, and helped me in times of need. I am honored
to have such an inspiring group of people that I am able to call my friends.
First, to Denia Djoki\'c, whose kindness knows no bounds and whose imagination is
unparalleled. Denia, I can't believe
that I met you on my first visit to Berkeley so many years ago, and we remain
friends to this day. To Katy Huff, who always has helped me
take pause and consider the deeper impact of my work and whose determination and
achievements are a continuous source inspiration for me.
To Alejandra Jolodosky, my cat-sister for life.
To Ashley Reichardt, who
helps me to see a different perspective every time we talk and who allows
me adventerous escapes from the day-to-day grind.
To Patricia Schuster, my
book club buddy and unwavering cheerleader. To Lakshana, for her
ability to be kind to any person in any circumstance,
even when they talk about her mom or change her background to animals kissing.
And to Kelly Rowland, my
commiserator in-chief. In no particular order,
thank you to my other academic friends: Nathan Bailey,
Sam Briggs, Sandra Bogetic, Tomi Akindele,
Anagha Iynegar, Perry Chodash, James Bevins, Seth Cadell,
and Micah Folsom.
Thanks to my past collaborators on non-dissertation research, in
particular Leah Morgan. Finally, thank you to the fellows and staff at the
Berkeley Institute for Data Science for allowing me to write and work on this
project in a stimulating, coffee-abundant, invigorating place.

Finally, I must thank my nuclear
family for sowing and nurturing my desire for
knowledge, my California family for
ensuring I lived a life outside of graduate
school, and my bay area friends for making the past several years in California
so special.
Last, I'd also like to thank Paul, for his unwavering
support, patience, and kindness. \\ \\

\footnotesize{This material is based on work supported by the Department of Energy under award
number DE- NE0008286. This report was prepared as an account of work sponsored
by an agency of the United States Government. Neither the United States
Government nor any agency thereof, nor any of their employees, makes any
warranty, express or implied, or assumes any legal liability or responsibility
for the accuracy, completeness, or usefulness of any information, apparatus,
product, or process disclosed, or represents that its use would not infringe
privately owned rights. Reference herein to any specific commercial product,
process, or service by trade name, trademark, manufacturer, or otherwise does
not necessarily constitute or imply its endorsement, recommendation, or favoring
by the United States Government or any agency thereof. The views and opinions of
the authors expressed herein do not necessarily state or reflect those of the
United States Government or any agency thereof.}

\end{acknowledgements}
